---
title: "Football Player Stats"
author: "Omar Khattab"
format: html
---

# Introduction

For this project, I used the **2022–2023 Football Player Stats** dataset. The dataset includes performance statistics for football players across European club teams.

The 3 questions:

1. Do Brazilian players score more goals or italian players.

^^Do South American players score significantly more goals than European players?
^^(This will be done by grouping players into 2 (Europeans) and (South Americans)

2. Is there a relationship between minutes played and goals scored?

3. Do more completed passed into the 18 yard box result in more assists?

# Data Description

Here are the key variables:

- `Name`: Player’s name  
- `Nation`: Country of the player  
- `Goals`: Number of goals scored  
- `MinutesPlayed`: Total minutes played  
- `Position`: Playing position (e.g., FW, MF, DF)  
- `Age`: Player's age
- `PPA` : Completed passes into the 18-yard box


```{r}
library(tidyverse)

players <- read_delim("2022-2023-Football-Player-Stats.csv", delim = ";")
glimpse(players)

```


Distribution of Goals by Player Position
```{r}
ggplot(players, aes(x = Pos, y = Goals)) +
  geom_boxplot(fill = "lightblue") +
  labs(
    title = "Distribution of Goals by Player Position",
    x = "Position",
    y = "Goals Scored"
  ) +
  theme_minimal()

```

Histogram of Number of Players from Argentina Brazil and Italy

```{r}
filtered_players <- players %>%
  filter(Nation %in% c("ARG", "BRA", "FRA"))

ggplot(filtered_players, aes(x = Nation)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Number of Players: Argentina, Brazil, and Italy",
       x = "Nation",
       y = "Player Count") +
  theme_minimal()
```


Relationship between Minutes played and Goals scored
```{r}
ggplot(players, aes(x =Min, y = Goals)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Relationship Between Minutes Played and Goals Scored",
    x = "Minutes Played",
    y = "Goals Scored"
  ) +
  theme_minimal()

attackers <- players %>%
  filter(Pos == "FW")  

# Correlation between minutes played and goals scored (for attackers)
ggplot(attackers, aes(x = Min, y = Goals)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Relationship Between Minutes Played and Goals Scored (Attackers)",
    x = "Minutes Played",
    y = "Goals Scored"
  ) +
  theme_minimal()

# Filtering only attackers who played at least 900 minutes
forwards <- players %>%
  filter(Pos == "FW", Min >= 900)

ggplot(forwards, aes(x = Min, y = Goals)) +
  geom_point(alpha = 0.6, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Relationship Between Minutes Played and Goals Scored (Attackers, Min 500 Mins)",
    x = "Minutes Played",
    y = "Goals Scored"
  ) +
  theme_minimal()

```


Running two sample T-test on brazilian players based on position
```{r}

players %>%
  filter(Nation == "BRA") %>%
  group_by(Pos) %>%
  summarise(
    mean_goals = mean(Goals, na.rm = TRUE),
    sd_goals = sd(Goals, na.rm = TRUE),
    min_goals = min(Goals, na.rm = TRUE),
    max_goals = max(Goals, na.rm = TRUE)
  )

```
1. Do Brazilian players score more goals or italian players.

```{r}

brazil_italy <- players %>%
  filter(Nation %in% c("BRA", "ITA"))

t.test(Goals ~ Nation, data = brazil_italy)


players <- players %>%
  mutate(Continent = case_when(
    Nation %in% c("BRA", "ARG", "URU", "COL", "CHI", "ECU", "PER", "PAR", "VEN", "BOL") ~ "South America",
    Nation %in% c("ITA", "GER", "FRA", "ESP", "POR", "ENG", "BEL", "NED", "CRO", "SWE",
                  "SUI", "DEN", "NOR", "POL", "AUT", "GRE", "SRB", "SCO", "UKR", "CZE") ~ "Europe",
    TRUE ~ "Other"
  ))


south_europe <- players %>%
  filter(Continent %in% c("South America", "Europe"))

t.test(Goals ~ Continent, data = south_europe)


```


Do more completed passes into 18 yard box result in more assists?
```{r}


ggplot(players, aes(x = PPA, y = Assists)) +
  geom_point(alpha = 0.6, color = "darkgreen") +
  geom_smooth(method = "lm", se = TRUE, color = "orange") +
  labs(
    title = "Relationship Between Passes Into Box and Assists",
    x = "Completed Passes Into 18-Yard Box",
    y = "Assists"
  ) +
  theme_minimal()

ggplot(players, aes(x = PPA, y = Assists)) +
  geom_point(alpha = 0.6, color = "darkgreen") +
  geom_smooth(method = "lm", se = TRUE, color = "orange") +
  labs(
    title = "Relationship Between Passes Into Box and Assists",
    x = "Completed Passes Into 18-Yard Box",
    y = "Assists"
  ) +
  theme_minimal()

players_filtered <- players %>%
  filter(PPA >= 1.2)

ggplot(players_filtered, aes(x = PPA, y = Assists)) +
  geom_point(alpha = 0.6, color = "purple") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Completed Passes Into Box vs Assists (PPA ≥ 0.3)",
    x = "Completed Passes Into 18-Yard Box (CrsPA)",
    y = "Assists"
  ) +
  theme_minimal()




```


Conclusion


I performed a two-sample t-test comparing the number of goals scored by Brazilian and Italian players. The p-value was 0.023, indicating a statistically significant difference.
On average, Brazilian players scored more goals (1.36) than Italian players (0.84).However when I tried with European and South American players it was very similar and not statistically significant (p=1.05)

There is a relationship between more minutes played and goals scored. The more you play the more likely you are to score atleast for attackers.

I could not really draw a result from the PPA and assists.

